---
output: word_document
editor_options: 
  chunk_output_type: console
---
#7TotaldescriptivaClean5.Rmd
#install.packages("rmarkdown")
#library("rmarkdown")

Set the folder where the word file is going to be generated
```{r}
dd<- read.csv("merged_data.csv")

```  

Check the type of the R object created for the dataset

```{r}            
class(dd)
```

Get dimensions of the dataset
````{r}
dim(dd)
n<-dim(dd)[1]
K<-dim(dd)[2]

n
K
```
Check the variables 

```{r}
names(dd)
```
Decide if you need to declare some more factor or date

```{r}
descriptiva<-function(X, nom){
  if (!(is.numeric(X) || class(X)=="Date")){ 
    frecs<-table(as.factor(X), useNA="ifany")
    proportions<-frecs/n
    
    pie(frecs, cex=0.6, main=paste("Pie of", nom))
    barplot(frecs, las=3, cex.names=0.7, main=paste("Barplot of", nom), col=listOfColors)
    print(paste("Number of modalities: ", length(frecs)))
    print("Frequency table")
    print(frecs)
    print("Relative frequency table (proportions)")
    print(proportions)
    print("Frequency table sorted")
    print(sort(frecs, decreasing=TRUE))
    print("Relative frequency table (proportions) sorted")
    print(sort(proportions, decreasing=TRUE))
   }else{
     if(class(X)=="Date"){
       print(summary(X))
       print(sd(X))
       #decide breaks: weeks, months, quarters...
       hist(X,breaks="weeks")
     }else{
       hist(X, main=paste("Histogram of", nom))
       boxplot(X, horizontal=TRUE, main=paste("Boxplot of",nom))
       print("Extended Summary Statistics")
       print(summary(X))
       print(paste("sd: ", sd(X, na.rm=TRUE)))
       print(paste("vc: ", sd(X, na.rm=TRUE)/mean(X, na.rm=TRUE)))
      }
   }
}

dataset<-dd
actives<-c(1:K)

```


Basic descriptive analysis for numerical variables

(decide the maximum number of colors you can need in a graph based on your metadata file)
```{r, echo=FALSE}
listOfColors<-rainbow(39)
# 
par(ask=TRUE)

for(k in actives){
  print(paste("variable ", k, ":", names(dd)[k] ))
  descriptiva(dd[,k], names(dd)[k])
}
par(ask=FALSE)

```


